\section{Conclusion}

This chapter has presented the readers with a comprehensive guide to some of the essential tools and techniques for data analysis and software development.
Understanding version control and utilizing Github enables one to manage their projects better, collaborate effectively with others and maintain a traceable history of the code. Utilizing IDEs such as R Studio or VS code offer numerous features that help streamline the workflow, improve productivity and facilitate debugging code and discovering errors more effectively.

The readers were introduced to using R and Python and were exposed to the usage of a subset of their vast number of libraries. We hope that by introducing some of the core concepts of each language, the reader is better equipped to start with data analysis and is better able to understand and explore the extensive Python and R ecosystems.

The end-to-end example, focused on analyzing simulated seizure spikes and count data, demonstrated how R and Python can be used to load, join, clean and visualize data to enable the extraction of meaningful insights from data which is stored in files with different formats and different folders. Further, best practices for creating modular functions and clean reusable code were presented.

It is our hope that this has set the foundation for the readers journey into programming and data analysis and we have equipped the readers with the ability to follow along the code presented in the subsequent chapters.
